\chapter*{ACKNOWLEDGMENTS}

I can definitely say that I never would have completed this thesis without overwhelming support, encouragement, and advisement from so many people, and I will do my best to thank everyone here. 

First, I have to thank my undergraduate research advisor Matthew Bellis. 
Thank you for encouraging me to join your group even when I thought doing research was beyond my capabilities, and then helping me realize it wasn't.
To my graduate advisor Jim Napolitano for being exactly the advisor that I needed. 
Thank you for guiding me from the timid first year student I was to the confident researcher I am now. 
What I appreciated most was your ability to provide me with constant support while also encouraging my independence.
To Don Jones, for his immense help throughout this whole process.
Thank you for taking the time to answer all of my questions and for providing valuable feedback at every step of my analysis.

Of course I must thank the other members of the Temple Team as well, Adam Hansell and James Wilhelmi.
Thank you for great lunch conversations and making my time at Temple enjoyable. 
To Adam, thank you for always listening to my rambles. I greatly enjoyed sharing an office space with you. 

I also would like to thank the PROSPECT collaboration as a whole. It has been a pleasure working with this great group of scientists. 
A specific thanks to David Jaffe, for his patience, expertise, and guidance throughout my entire project. 
And thank you to Danielle Noricini for making me feel welcome from the beginning, and for being a great leader during the construction process. 

I am extremely grateful to all my friends and family that have stuck by me through this whole process.
To Shannon, Aneline, and Meghan - it is a privilege to be surrounded by successful women and to be able to call you my best friends.
To the Landschoots, the family I officially gained while on this journey, thank you supporting me and cheering me on. 
Thank you to everyone in the Berish-Krentler clan. Our holidays and summer vacations always gave me something to look forward to, and re-energized my return to work.
A huge thank you to Grandma. Thank you for, without a doubt, being my biggest cheerleader from day one. 
To Kristen, Michael, and Robert, I could not have asked for better siblings, and I appreciated your support along the way - even if you still are not entirely sure what I did over the past five-and-a-half years.

To Ben, my husband, thank you for being by my side through it all. 
Thank you for encouraging me, supporting me, pushing me to finish, and picking up the slack around the house when I got too busy.
Your support was essential for me to complete this thesis.

Lastly, and most importantly, I have to thank my parents. 
None of this would have been possible without all that you have done for me. 
Your constant support and encouragement have been invaluable, and I am forever grateful to you. 
\\ \\
\begin{center}
\textit{``Faith and reason are like two wings on which the human spirit rises to the contemplation of truth."} - Pope John Paul II
\end{center}
